hbm_paper_results

### 2024-01-22 paper_prop_results.Rmd
prep to rerun featurewise sex tests

```{r}
data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"

@@

#also load raw data
ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")

@@

# Secondary: Vary M:F

@@

Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)

@@

```{r}
ratio.df %>%
  dplyr::filter(dataset=="cf") %>%
ggplot() +
  geom_bar(aes(x=prop, fill=sex)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ sim.site)
```

@@

Do centile score errors change based on combat version?
*need to look and see what happens for subjects who aren't in all samples?*

```{r}
prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
names(prop_abs.cent_t.tests) <- n_prop_list
prop_abs.cent_t.tests_df <- bind_rows(prop_abs.cent_t.tests, .id = "column_label")
```

Visualize 
```{r}
ggplot(prop_abs.cent_t.tests_df) +
  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
  geom_hline(yintercept = 208, linetype='longdash') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
  facet_wrap(~comp)
```

@@

prop.global_cent_errs_plot <- prop.global_abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
         abs.diff = abs.diff*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Study Site") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels=percent_f) +
  facet_wrap(~ pheno) +
  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
prop.global_cent_errs_plot
#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
```

@@

>>>>

```{r}
der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"

@@

@@

@@

Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
```{r, eval=FALSE}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)

@@

```{r, eval=FALSE}
ratio.df %>%
  dplyr::filter(dataset=="cf") %>%
ggplot() +
  geom_bar(aes(x=prop, fill=sex)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ sim.site)
```

@@

Do centile score absolute errors change based on combat version? Tested by `ratio_results.R`, results written to `full_featurewise_cent_t_tests.csv`
*need to look and see what happens for subjects who aren't in all samples?*

```{r}
cent.t.test.df <- fread(paste0(data_path, "/ukb_ratio/full_featurewise_cent_t_tests.csv"))
cent.t.test.df <- cent.t.test.df %>%
  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
         comp = as.factor(comp))
```

```{r}
cent.t.test.df %>%
  # filter(comp=="cf.gamlss_cf.gam") %>%
  ggplot() +
  geom_col(aes(x=prop, y=n_sig_regions, fill=comp), position="dodge") +
  facet_wrap(~comp) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
```

Plot errors for global phenotypes
```{r}
percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))

@@

prop.global_cent_errs_plot <- prop.global_abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
         abs.diff = abs.diff*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Study Site") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels=percent_f) +
  facet_wrap(~ pheno) +
  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
prop.global_cent_errs_plot
#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
```
*need to add stat tests to plot*

@@

Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_cent_t_tests.csv`

```{r}
cent.subj.mean.test.df <- fread(paste0(data_path, "/ukb_ratio/subj.abs.mean_sex_bias_cent_t_tests.csv"))
cent.subj.mean.test.df <- cent.subj.mean.test.df %>%
  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
         dataset = as.factor(dataset))
cent.subj.mean.test.df
```

```{r}
cent.subj.mean.test.df %>%
  # filter(comp=="cf.gamlss_cf.gam") %>%
  ggplot() +
  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
```

Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`

```{r}
cent.subj.sex.test.list <- readRDS(paste0(data_path, "/ukb_ratio/full_featurewise_cent_sex_bias_tests.RDS"))

```
